---
title: "FinalProjectAnalysis"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, results='hide'}
library(lme4)
## Load Data and EDA/Check Assumptions
sleepDat<-sleepstudy
hist(sleepDat$Reaction)
hist(sleepDat$Days)

sleepMod1<-lm(Reaction~Days + Subject, sleepDat)
par(mfrow=c(2,2))
plot(sleepMod1)

library(gvlma)
gvlma(sleepMod1)
library(car)
influenceIndexPlot(sleepMod1)
vif(sleepMod1)

par(mfrow=c(1,1))

library(ggplot2)
ggplot(sleepDat, aes(x = Days, y = Reaction, colour = Subject)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(legend.position = "none")

## Possible Influence/outliers, independence assumption is violated, homoscedasity may be violated. Going to try linear mixed effects models to see if they will fix the assumptions.
```

```{r, results='hide'}
DayCol<-sleepDat$Days
ReactCol<-sleepDat$Reaction
plot(DayCol, ReactCol)
## Now build linear mixed effects models
## Random Intercept first
sleepMod2<-lmer(Reaction ~ Days + (1|Subject), sleepDat)
summary(sleepMod2)
coef(sleepMod2)


## Random Slope Model
sleepMod3<-lmer(Reaction ~ Days + (Days|Subject), sleepDat)
summary(sleepMod3)
coef(sleepMod3)



## Plotting Mixed Effects Models
## Random Intercept Model
fixParam<-fixef(sleepMod2)

randParamVector<-coef(sleepMod2)[["Subject"]][["(Intercept)"]]
SubjectVector<-levels(sleepDat$Subject)

plot(Reaction~Days,data=sleepDat,ylab="Reaction Time",xlab="Days", ylim=c(100, 500), main="Random Intercept Model of Reaction Time vs. Days \nby Subject")
subNum<-unique(sleepDat$Subject)
for(i in 1:18){
abline(a=randParamVector[i], b=fixParam[2],col=(20+i),lty=2,lwd=2)

}
abline(fixParam,lwd=6,col="red")
legend("bottomright", legend=SubjectVector,
       col=21:38, lty=2, cex=0.375, horiz=TRUE)


## Random Slope Model
fixParam<-fixef(sleepMod3)

randParamVector<-coef(sleepMod3)[["Subject"]][["(Intercept)"]]
fixParamVector<-coef(sleepMod3)[["Subject"]][["Days"]]

plot(Reaction~Days,data=sleepDat,ylab="Reaction Time",xlab="Days", ylim=c(100, 500), main="Random Slope and Intercept Model of Reaction Time vs. Days \nby Subject")
subNum<-unique(sleepDat$Subject)
for(i in 1:18){
abline(a=randParamVector[i],b=fixParamVector[i], col=(20+i), lty=2, lwd=2)

}
abline(fixParam,lwd=6,col="red")
legend("bottomright", legend=SubjectVector,
       col=21:38, lty=2, cex=0.375, horiz=TRUE)


## Likelihood Ratio Test
##Random Intercept Model Fixed Effect Significance
sleepMod2<-lmer(Reaction ~ Days + (1|Subject), sleepDat, REML = FALSE)
sleepMod2.Null<-lmer(Reaction ~ 1 + (1|Subject),sleepDat,REML = FALSE)

anova(sleepMod2.Null, sleepMod2)
## Chi-Sq:116.46, Df=1, p-value:2.2e-16
## Days is significant for the model

## Likelihood Ratio Test
## Fixed Effect Model Random effect Significance
sleepMod2<-lmer(Reaction ~ Days + (1|Subject), sleepDat)
library(RLRsim)
exactRLRT(sleepMod2)

## P-Value <2.2e-16 is evidence the random effect is non zero

## Likelihood Ratio Test
## Random Slope Model
sleepMod3<-lmer(Reaction ~ Days + (Days|Subject), sleepDat, REML = FALSE)
sleepMod3.Null<-lmer(Reaction~1 + (Days|Subject), sleepDat, REML=FALSE)
sleepMod3.RanNull<-lm(Reaction~Days, sleepDat)

anova(sleepMod3.Null, sleepMod3)
## Chi-Square: 23.537, Df=1, P-Value:1.226e-06
## Days is significant in the model

## Restricted Likelihood Ratio Test
## Random Slope Effect test for significance

sleepMod2<-lmer(Reaction ~ Days + (1|Subject), sleepDat)
mA <- update(sleepMod2, .~. + (0 + Days|Subject))
mSlope <- update(mA, .~. - (1|Subject))
exactRLRT(mSlope, mA, m0=sleepMod2)



## Checking assumptions
plot(sleepMod1)
library(car)
influenceIndexPlot(sleepMod1)

plot(sleepMod2)
qqnorm(resid(sleepMod2))
qqline(resid(sleepMod2))

plot(sleepMod3)
qqnorm(resid(sleepMod3))
qqline(resid(sleepMod3))

